fl0w - a development environment for the\\ KIPR Wallaby

### 2017-04-01 paper.tex
Added citation to disc0very figure
(Hopefully) final changes

\documentclass[conference]{IEEEtran}

@@

\begin{abstract}
This publication introduces fl0w\cite{fl0w:Philip Trauner}, an alternative\\ development environment for the KIPR Wallaby\cite{Wallaby Controller:KIPR}.\\ The aim of fl0w\cite{fl0w:Philip Trauner} is to improve the Botball program development experience. It focuses on the components that make up fl0w\cite{fl0w:Philip Trauner}, namely a file synchronization protocol that maintains a consistent state across all connected controllers, a route based and data type preserving network protocol with peer-to-peer piping capabilities, a discovery protocol that connects all controllers together automatically, a Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin which enables in-line sensor readouts, program execution, program editing, and keyboard shortcuts, and a browser-based management front-end to manage the controller fleet.\\
\end{abstract}

\begin{IEEEkeywords}
file synchronization, networking, LAN discovery, development environment
\end{IEEEkeywords}

@@

\section{Implementation}
All fl0w\cite{fl0w:Philip Trauner} components emphasize shared code and are therefor written in Python\cite{Python:Python Foundation} 3.3.6 to remain compatible with the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin environment. \\Python\cite{Python:Python Foundation} was chosen as the implementation language for fl0w\cite{fl0w:Philip Trauner} because of its rich standard library and its lightweight virtual machine that can run on a Wallaby\cite{Wallaby Controller:KIPR} controller with acceptable speed. A pre-compiled version of Python\cite{Python:Python Foundation} is bundled with the fl0w\cite{fl0w:Philip Trauner} installer, r0adrunner, because it is not present on the Wallaby\cite{Wallaby Controller:KIPR} by default. \\All fl0w\cite{fl0w:Philip Trauner} components are designed as libraries to allow for integration into other projects. fl0w\cite{fl0w:Philip Trauner} utilities a client-server networking model with random server assignment. Client-server was chosen because the server can act as a data source for all clients.

@@

The Python\cite{Python:Python Foundation} version is built on top of a fork of ws4py\cite{ws4py:Philip Trauner} and the JavaScript version utilizes regular WebSockets\cite{The WebSocket Protocol:A. Melnikov} in arraybuffer mode. It was created to mitigate the problems fl0ws\cite{fl0w:Philip Trauner}explorative development style created. Instead of one monolithic network protocol, undergr0und emphasizes small sub protocols. This was done because all monolithic designs became too inconsistent over time. undergr0und\cite{undergr0und:Philip Trauner} manages itself through its own concepts.\\

\subsubsection{undergr0und.js}
undergr0und.js\cite{undergr0und.js:Philip Trauner} is the feature complete client-side JavaScript port of undergr0und\cite{undergr0und:Philip Trauner} that was required for dashb0ard\cite{dashb0ard:Sebastian Schaffler}, the browser-based front-end for fl0w\cite{fl0w:Philip Trauner}. \\It utilizes node-jspack\cite{node-jspack:Peter Griess} to unpack the binary messages it receives from the undergr0und\cite{undergr0und:Philip Trauner} server. It is based on the Python\cite{Python:Python Foundation} version but does not implement the server because there is currently no need for that functionality. \\undergr0und.js can run be used in Node.js\cite{Node.js:Node.js Foundation} as well as browsers that support WebSockets\cite{The WebSocket Protocol:A. Melnikov}. browserify\cite{browserify:James Halliday} is used to create the browser version.\\

@@

\subsubsection{Data type preservation}
The original type of the data segment is present inside the header $($see Table ~\ref{fig:undergr0und_header}$)$ to accurately reconstruct sent data on the other end. JSON\cite{JSON:T. Bray Ed.} is used to transfer lists and dictionaries, and regular ASCII encoded strings are utilized to transfer integers, floats, null/none types and strings.

@@

\subsubsection{Synchronization}
On startup, behem0th\cite{behem0th:Christoph Heiss} synchronizes files based on their last modification time and MD5 hash. After that, files get synchronized to other clients as soon as a file-system event happens. Synchronization conflicts are resolved on the server and it is possible to use behem0th\cite{behem0th:Christoph Heiss} with a theoretically indefinite amount of clients.\\

@@

\subsubsection{Protocol}
behem0th\cite{behem0th:Christoph Heiss} is independent from underg0und and defines its own network protocol. It is completely text-based, uses UTF-8\cite{UTF-8:F. Yergeau} for character encoding and JSON\cite{JSON:T. Bray Ed.} as format encoding. behem0th\cite{behem0th:Christoph Heiss} is designed around 'requests' $($see Figure ~\ref{fig:behem0th_request}$)$ and 'routes'. Each request is a JSON-formatted string and separated by a newline.

@@

\subsection{dashb0ard}
dashb0ard\cite{dashb0ard:Sebastian Schaffler} is a single-page web front-end designed to manage the fl0w\cite{fl0w:Philip Trauner} fleet.\\ It can be used to configure connected controllers and obtain sensor readouts as well as debug logs. \\It utilities the JavaScript version of undergr0und to retrieve data from fl0w\cite{fl0w:Philip Trauner}, Vue.js\cite{Vue:Evan You} for its user interface, Bootstrap 3\cite{Bootstrap 3:Twitter Inc.} as a design baseline, Flot\cite{Flot:David Schnur} to display graphs and jQuery\cite{jQuery:jQuery Foundation} for additional DOM manipulation.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{dashb0ard-sensor}
\caption{The sensor readout page of dashb0ard\cite{Sublime Text 3:Sublime HQ}}
\label{fig:dashb0ard_sensor}
\end{figure}
\subsection{edit0r}
edit0r\cite{edit0r:Philip Trauner} is the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin. It is modeled as a fl0w\cite{fl0w:Philip Trauner} client and allows for remote robot management, source code synchronization as well as in-line sensor readouts. \\It also provides quick access to dashb0ard\cite{dashb0ard:Sebastian Schaffler}, which is required because the IP addresses of controllers are not always static. behem0th\cite{behem0th:Christoph Heiss} is embedded into edit0r\cite{edit0r:Philip Trauner} and enables the source code synchronization. In its current form only C programs are supported.\\

@@

\subsubsection{In-line sensor readouts}
Invocations of the functions "analog" and "digital" are located every time the content of a view changes and the parameters of the calls are grouped across views. edit0r\cite{edit0r:Philip Trauner} subscribes to all required sensor ports by requesting them from the currently selected controller. The controller keeps track of all subscriptions to sensors and continuously transmits the appropriate values to the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin. \\Sublime Text 3 \cite{Sublime Text 3:Sublime HQ} phantoms are created as soon as new sensor values are available. This approach has the disadvantage of higher than usual processor utilization by edit0r\cite{edit0r:Philip Trauner}, because phantoms were primarily designed to display infrequently updated content. To partially circumvent this, sensor readouts can be disabled and are not enabled by default.\\
In-line sensor readouts can be toggled on a per view basis with a keyboard shortcut (default F10).\\

@@

\subsection{r0adrunner}
r0adrunner is the installer of fl0w\cite{fl0w:Philip Trauner}. \\It is built to mimic a Wallaby\cite{Wallaby Controller:KIPR} controller software update to remain compatible with the preexisting update functionality found in Harrogate\cite{Harrogate:KIPR}. It is implemented in Python\cite{Python:Python Foundation} 2.7, which is already installed on all Wallaby\cite{Wallaby Controller:KIPR} controllers but because fl0w\cite{fl0w:Philip Trauner} utilizes features of Python\cite{Python:Python Foundation} 3.3 it installs a full-fledged precompiled version of Python\cite{Python:Python Foundation} 3.6.0. \\It also changes the Wallaby\cite{Wallaby Controller:KIPR} hostname to their unique manufacturing ID because fl0w\cite{fl0w:Philip Trauner} utilizes hostnames to represent controllers in its user facing components. r0adrunner is distributed as an archive that has to be copied to an USB storage medium.

\subsection{disc0very}
disc0very\cite{disc0very:Christoph Heiss} is the LAN discovery protocol of fl0w\cite{fl0w:Philip Trauner}. It relies on UDP\cite{UDP:J. Postel} broadcasts to coordinate server and client assignment. Wallaby controller instances as well as the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin edit0r\cite{edit0r:Philip Trauner} utilize it to determine if there is already an active fl0w server on the network upon startup without user interaction.

\subsubsection{Discovery process}
Upon launch disc0very\cite{disc0very:Christoph Heiss} listens for advertisements from an already running fl0w\cite{fl0w:Philip Trauner} server. If an advertisement is received the client-mode is engaged and a connection to the server is established. If no advertisement has been received after 2000ms server-mode is engaged. Server-mode can be disabled for clients that do not have the capability to become a server (for example edit0r\cite{edit0r:Philip Trauner}).

\begin{figure}[H]
\centering
\begin{json}
{
    "address": "<address>",
    "port": "<port>"
}
\end{json}
\caption{disc0very advertisement encoding in JSON\cite{JSON:T. Bray Ed.}}
\label{fig:discovery_advertisement}
\end{figure}

@@

\subsubsection{Server}
The server keeps track of all clients and provides means to acquire their peer IDs, which are used by undergr0und to target specific clients. Additionally it serves dashb0ard\cite{dashb0ard:Sebastian Schaffler} and handles program compilation. It utilizes undergr0und for networking, behem0th for file-synchronization, and disc0very to determine if another server is already running in the network. \\
In that situation the start is interrupted, the Wallaby\cite{Wallaby Controller:KIPR} client is started and it connects to the available server instance, otherwise server and client are started. \\

\section{Conclusion}

After prolonged testing it was deemed reasonable to claim that fl0w\cite{fl0w:Philip Trauner} can improve the Botball development work-flow on Wallaby\cite{Wallaby Controller:KIPR} controllers.

@@

Many edge cases can and will occur while performing continuous file-synchronization with an undetermined amount of clients across multiple operating-systems and file-systems.

@@

% section* for acknowledgment
\section*{Acknowledgment}
The authors would like to thank the robotics team robot0nfire: Nico Kratky, Nico Leidenfrost, Sebastian Schaffler, Christine Zeh, Sascha Zemann;\\ Dr. Michael Stifter for making the existence of our team possible; Daniel Maximilian Swoboda for answering all paper related questions; the KIPR development team without whom the Wallaby\cite{Wallaby Controller:KIPR} Controller would not exist.

@@

>>>>

\documentclass[conference,a4paper]{IEEEtran}

@@

\begin{abstract}
This publication introduces fl0w, an alternative development environment and monitoring solution for the KIPR Wallaby. The aim of fl0w is to improve the Botball program development experience. It focuses on the components that make up fl0w, namely a file synchronization protocol that maintains a consistent state across all connected controllers, a route based and data type preserving network protocol with peer-to-peer piping capabilities, a discovery protocol that connects all controllers together automatically, a Sublime Text 3 plugin which enables in-line sensor readouts, program execution, program editing, and keyboard shortcuts, and a browser-based management front-end to manage the controller fleet.\\
\end{abstract}

\begin{IEEEkeywords}
file synchronization, networking, LAN discovery, development environment, monitoring
\end{IEEEkeywords}

@@

\section{Implementation}
All fl0w\cite{fl0w:Philip Trauner} components emphasize shared code and are therefor written in Python\cite{Python:Python Foundation} 3.3.6 to remain compatible with the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin environment. \\Python was chosen as the implementation language for fl0w because of its rich standard library and its lightweight virtual machine that can run on a Wallaby\cite{Wallaby Controller:KIPR} controller with acceptable speed. A pre-compiled version of Python is bundled with the fl0w installer, r0adrunner, because it is not present on the Wallaby by default. \\All fl0w components are designed as libraries to allow for integration into other projects. fl0w utilities a client-server networking model with random server assignment. Client-server was chosen because the server can act as a data source for all clients.

@@

The Python\cite{Python:Python Foundation} version is built on top of a fork of ws4py\cite{ws4py:Philip Trauner} and the JavaScript\cite{ECMAScript:Ecma} version utilizes regular WebSockets\cite{The WebSocket Protocol:A. Melnikov} in arraybuffer mode. It was created because the requirements for fl0ws\cite{fl0w:Philip Trauner} network protocol specification changed constantly and a dynamic approach to networking was required. Instead of one monolithic network protocol, undergr0und emphasizes small sub protocols. undergr0und manages itself through its own concepts.\\

\subsubsection{undergr0und.js}
undergr0und.js\cite{undergr0und.js:Philip Trauner} is the feature complete client-side JavaScript\cite{ECMAScript:Ecma} port of undergr0und\cite{undergr0und:Philip Trauner} that was required for dashb0ard\cite{dashb0ard:Sebastian Schaffler}, the browser-based front-end for fl0w\cite{fl0w:Philip Trauner}. \\It utilizes node-jspack\cite{node-jspack:Peter Griess} to unpack the binary messages it receives from the undergr0und server. It is based on the Python\cite{Python:Python Foundation} version but does not implement the server because there is currently no need for that functionality. \\undergr0und.js can be used in Node.js\cite{Node.js:Node.js Foundation} as well as browsers that support WebSockets\cite{The WebSocket Protocol:A. Melnikov}. browserify\cite{browserify:James Halliday} is used to create the browser version.\\

@@

\subsubsection{Data type preservation}
The original type of the data segment is present inside the header $($see Table ~\ref{fig:undergr0und_header}$)$ to accurately reconstruct sent data on the other end. JSON\cite{JSON:T. Bray Ed.} is used to transfer lists and dictionaries, and regular ASCII encoded strings are utilized to transfer integers, floats, null/none types and strings. JSON was chosen because it can be serialized and parsed by Python\cite{Python:Python Foundation} as well as JavaScript\cite{ECMAScript:Ecma}.

@@

\subsubsection{Synchronization}
On startup, behem0th\cite{behem0th:Christoph Heiss} synchronizes files based on their last modification time and MD5 hash. After that, files get synchronized to other clients as soon as a file-system event happens. Synchronization conflicts are resolved on the server and it is possible to use behem0th with a theoretically indefinite amount of clients.\\

@@

\subsubsection{Protocol}
behem0th\cite{behem0th:Christoph Heiss} is independent from underg0und and defines its own network protocol. It is completely text-based, uses UTF-8\cite{UTF-8:F. Yergeau} for character encoding and JSON\cite{JSON:T. Bray Ed.} as format encoding. behem0th is designed around 'requests' $($see Figure ~\ref{fig:behem0th_request}$)$ and 'routes'. Each request is a JSON-formatted string and separated by a newline.

@@

\subsection{dashb0ard}
dashb0ard\cite{dashb0ard:Sebastian Schaffler} is a single-page web front-end designed to manage the fl0w\cite{fl0w:Philip Trauner} fleet.\\ It can be used to configure connected controllers and obtain sensor readouts as well as debug logs. \\It utilities undergr0und.js\cite{undergr0und.js:Philip Trauner} to retrieve data from fl0w, Vue.js\cite{Vue:Evan You} for its user interface, Bootstrap 3\cite{Bootstrap 3:Twitter Inc.} as a design baseline, Flot\cite{Flot:David Schnur} to display graphs and jQuery\cite{jQuery:jQuery Foundation} for additional DOM manipulation.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{dashb0ard-sensor}
\caption{The sensor readout page of dashb0ard\cite{Sublime Text 3:Sublime HQ}}
\label{fig:dashb0ard_sensor}
\end{figure}
\subsection{edit0r}
edit0r\cite{edit0r:Philip Trauner} is the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin. It is modeled as a fl0w\cite{fl0w:Philip Trauner} client and allows for remote robot management, source code synchronization as well as in-line sensor readouts. \\It also provides quick access to dashb0ard\cite{dashb0ard:Sebastian Schaffler}, which is required because the IP addresses of controllers are not always static. behem0th\cite{behem0th:Christoph Heiss} is embedded into edit0r and enables the source code synchronization. In its current form only C programs are supported.\\

@@

\subsubsection{In-line sensor readouts}
Invocations of the functions "analog" and "digital" are located every time the content of a view changes and the parameters of the calls are grouped across views. edit0r\cite{edit0r:Philip Trauner} subscribes to all required sensor ports by requesting them from the currently selected controller. The controller keeps track of all subscriptions to sensors and continuously transmits the appropriate values to the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin. \\Sublime Text 3 phantoms are created as soon as new sensor values are available. This approach has the disadvantage of higher than usual processor utilization by edit0r, because phantoms were primarily designed to display infrequently updating content. To partially circumvent this, sensor readouts can be disabled and are not enabled by default.\\
In-line sensor readouts can be toggled on a per view basis with a keyboard shortcut (default F10).\\

@@

\subsection{r0adrunner}
r0adrunner is the installer of fl0w\cite{fl0w:Philip Trauner}. \\It is built to mimic a Wallaby\cite{Wallaby Controller:KIPR} controller software update to remain compatible with the preexisting update functionality found in Harrogate\cite{Harrogate:KIPR}. It is implemented in Python\cite{Python:Python Foundation} 2.7, which is already installed on all Wallaby controllers. A full-fledged precompiled version of Python 3.6.0 is installed by r0adrunner because fl0w utilizes features of Python 3.3.6. \\It also changes the Wallaby hostname to their unique manufacturing ID because fl0w uses hostnames to represent controllers in its user facing components. r0adrunner is distributed as an archive that has to be copied to an USB storage medium.

\subsection{disc0very}
disc0very\cite{disc0very:Christoph Heiss} is the LAN discovery protocol of fl0w\cite{fl0w:Philip Trauner}. It relies on UDP\cite{UDP:J. Postel} broadcasts to coordinate server and client assignment. Wallaby controller instances as well as the Sublime Text 3\cite{Sublime Text 3:Sublime HQ} plugin edit0r\cite{edit0r:Philip Trauner} make use of it to determine if there is already an active fl0w server on the network upon startup without user interaction.

\subsubsection{Discovery process}
Upon launch disc0very\cite{disc0very:Christoph Heiss} listens for advertisements from an already running fl0w\cite{fl0w:Philip Trauner} server. If an advertisement is received the client-mode is engaged and a connection to the server is established. If no advertisement has been received after 2s server-mode is engaged. Server-mode can be disabled for clients that do not have the capability to become a server, for example edit0r\cite{edit0r:Philip Trauner}).

\begin{figure}[H]
\centering
\begin{json}
{
    "address": "<address>",
    "port": "<port>"
}
\end{json}
\caption{disc0very\cite{disc0very:Christoph Heiss} advertisement encoding in JSON\cite{JSON:T. Bray Ed.}}
\label{fig:discovery_advertisement}
\end{figure}

@@

\subsubsection{Server}
The server keeps track of all clients and provides means to acquire their peer IDs, which are used by undergr0und to target specific clients. Additionally it serves dashb0ard\cite{dashb0ard:Sebastian Schaffler} and handles program compilation. It utilizes undergr0und for networking, behem0th for file-synchronization, and disc0very\cite{disc0very:Christoph Heiss} to determine if another server is already running in the network. \\
In that situation the start is interrupted, the Wallaby\cite{Wallaby Controller:KIPR} client is started and it connects to the available server instance, otherwise server and client are started. \\

\section{Conclusion}
After prolonged testing it was deemed reasonable to claim that fl0w\cite{fl0w:Philip Trauner} can improve the Botball development work-flow on Wallaby\cite{Wallaby Controller:KIPR} controllers. Experienced Botball teams that would rather program in native tools than the browser-based Harrogate and want to observe their sensor-readouts in realtime can benefit from fl0w.

@@

There is a high amount of complexity involved when performing continuous file-synchronization with an undetermined amount of clients across multiple operating-systems and file-systems.

@@

\section*{Acknowledgment}
The authors would like to thank the robotics team robot0nfire: Nico Kratky, Nico Leidenfrost, Sebastian Schaffler, Christine Zeh, Sascha Zemann;\\ Dr. Michael Stifter for making the existence of our team possible; Daniel Maximilian Swoboda for answering all paper related questions; the KIPR development team without whom the Wallaby\cite{Wallaby Controller:KIPR} Controller would not exist.

@@

\bibitem{ECMAScript:Ecma}
Ecma International, \emph{ECMAScript}, \url{https://www.ecma-international.org/ecma-262/7.0/index.html},\\ language specification,
accessed February 12th 2017
